|

### 2021-12-17 01-Chp-Background.Rmd
Submitted version

<!----------------------------------------------------------------------------->
(ref:diagnosticCriteria-caption) __Overview of the DSM-5 criteria for dementia (major neurocognitive disorder).__ - These published criteria[@edition2013] are commonly used to define dementia cases. Criterion A captures the cognitive decline aspect of dementia, while Criteria B-D ensure that the impairment observed is sufficient to interfere with daily living and cannot explained by other causes. Note that all criteria (A-D) must be met.

>>>>

<!----------------------------------------------------------------------------->
(ref:diagnosticCriteria-caption) __Overview of the DSM-5 criteria for dementia (major neurocognitive disorder)__ - These published criteria[@edition2013] are commonly used to define dementia cases. Criterion A captures the cognitive decline aspect of dementia, while Criteria B-D ensure that the impairment observed is sufficient to interfere with daily living and cannot explained by other causes. Note that all criteria (A-D) must be met.

### 2021-12-17 03-Chp-Systematic-Review-Methods.Rmd
Submitted version

(ref:sys-rev-methods-quote) --- Archibald Cochrane, 2000 [@cochrane1979]

@@

Rater reliability was assessed using Gwet's agreement coefficient ($AC1$).[@gwet2008] This approach was chosen over other measures such as percent agreement because it accounts for chance agreement between reviewers but does not suffer from bias due to severely imbalanced marginal totals in the same way that Cohen's $kappa$ value does. [@cohen1960: @gwet2008; @wongpakaran2013] Given the small number of included studies in this review as a proportion of the total number screened, this is a useful characteristic. The formulae used to calculate $AC$ are given in Appendix \@ref(appendix-gwet).

@@

<!----------------------------------------------------------------------------->
(ref:gwet-caption) __Ranges for Gwet's AC1__ - To aid the interpretation of inter-/intra-rater reliability, I used a published set of suggested categories for Gwet's AC1.[@mchugh2012]

(ref:gwet-scaption) Ranges for Gwet's AC1

@@

Intra- and inter-rater reliability was assessed against these cut-offs. If this assessment demonstrated an issue with the screening process (defined as an $AC$ of less than 0.9), it indicated the need for a larger proportion of records to be dual screened.

>>>>

(ref:sys-rev-methods-quote) --- Archibald Cochrane, 2000[@cochrane1979]

@@

Rater reliability was assessed using Gwet's agreement coefficient ($AC1$).[@gwet2008] This approach was chosen over other measures such as percent agreement because it accounts for chance agreement between reviewers but does not suffer from bias due to severely imbalanced marginal totals in the same way that Cohen's $kappa$ value does. [@cohen1960: @gwet2008; @wongpakaran2013] Given the small number of included studies in this review as a proportion of the total number screened, this is a useful characteristic. The formulae used to calculate $AC1$ are given in Appendix \@ref(appendix-gwet).

@@

<!----------------------------------------------------------------------------->
(ref:gwet-caption) __Ranges for Gwet's $AC1$__ - To aid the interpretation of inter-/intra-rater reliability, I used a published set of suggested categories for Gwet's AC1.[@mchugh2012]

(ref:gwet-scaption) Ranges for Gwet's $AC1$

@@

Intra- and inter-rater reliability was assessed against these cut-offs. If this assessment demonstrated an issue with the screening process (defined as an $AC1$ of less than 0.9), it indicated the need for a larger proportion of records to be dual screened.

### 2021-12-17 04-Chp-Systematic-Review-Results.Rmd
Submitted version

---
bibliography: bibliography/references.bib
csl: bibliography/nature.csl
output:
  bookdown::pdf_document2:
    template: templates/brief_template.tex
  bookdown::word_document2: 
      toc: false
      toc_depth: 3
      reference_docx: templates/word-styles-reference-01.docx
      number_sections: false 
  bookdown::html_document2: default
documentclass: book
---
```{block type='savequote', include=knitr::is_latex_output(),quote_author='(ref:sys-rev-results-quote)', echo = TRUE}
Hold  
```

(ref:sys-rev-results-quote) --- Hold

@@

<!----------------------------------------------------------------------------->
(ref:obsStatinAD-cap) __Random-effects meta-analysis of statins on Alzheimer's disease__ Non-randomised studies examining the association of statin use with Alzheimer's disease were synthesised using a random-effects meta-analysis. Subgroup estimates, stratified by overall risk-of-bias level, are also presented. D1-D7 and O refer to the seven risk-of-bias domains and single overall judgement in the ROBINS-I tool, while the risk of bias due to missing evidence is shown beside the overall summary diamond.

>>>>

---
bibliography: bibliography/references.bib
csl: bibliography/nature.csl
output:
  bookdown::pdf_document2:
    template: templates/brief_template.tex
  bookdown::word_document2: 
      toc: false
      toc_depth: 3
      reference_docx: templates/word-styles-reference-01.docx
      number_sections: false 
  bookdown::html_document2: default
documentclass: book
---
```{block type='savequote', include=knitr::is_latex_output(),quote_author='(ref:sys-rev-results-quote)', echo = TRUE}
The hundreds of hours spent conducting a scientific study ultimately contribute only a piece of an enormous puzzle. The value of any single study is derived from how it fits with and expands previous work, as well as from the study's intrinsic properties. Through systematic review the puzzle's intricacies may be disentangled.  
```

(ref:sys-rev-results-quote) --- Cynthia D Mulrow, 1994[@mulrow1994]

@@

<!----------------------------------------------------------------------------->
(ref:obsStatinAD-cap) __Random-effects meta-analysis of statins on Alzheimer's disease__ - Non-randomised studies examining the association of statin use with Alzheimer's disease were synthesised using a random-effects meta-analysis. Subgroup estimates, stratified by overall risk-of-bias level, are also presented. D1-D7 and O refer to the seven risk-of-bias domains and single overall judgement in the ROBINS-I tool, while the risk of bias due to missing evidence is shown beside the overall summary diamond.

### 2021-12-17 05-Chp-CPRD-Analysis.Rmd
Submitted version

<!----------------------------------------------------------------------------->
(ref:controlOutcomeFig-cap) __Sensitivity analysis: control outcomes__ Effect estimates obtained from the fully-adjusted time-varying model with participant age as the time scale for three control outcomes: backpain (negative), ischemic heart disease (positive protective), and diabetes (positive harmful).

>>>>

<!----------------------------------------------------------------------------->
(ref:controlOutcomeFig-cap) __Sensitivity analysis: control outcomes__ -  Effect estimates obtained from the fully-adjusted time-varying model with participant age as the time scale for three control outcomes: backpain (negative), ischemic heart disease (positive protective), and diabetes (positive harmful).

### 2021-12-17 07-Chp-Triangulation.Rmd
Submitted version

```{block type='savequote', include=knitr::is_latex_output(), quote_author='(ref:tri-quote)', echo = TRUE}
Next on my list of features to be specially considered I would place the consistency of the observed association. Has it been repeatedly observed by different persons, in different places, circumstances and times?
```
(ref:tri-quote) --- Austin Bradford Hill, 1965 [@hill1965]

@@

There are two key outputs from this chapter. The first is an R package, `triangulate`. Personal communication with the authors of the bias-/indirectness-adjusted meta-analysis approach resulted in the STATA code to implement the adjusted model.[@turner2009] This has since been generalised as part of the `triangulate` R package to enable other users to apply the approach detailed here. The package allows for preprocessing of domain-based risk-of-bias data to correctly identify the direction of bias relative to the effect estimate, use of domain-specific prior distributions, production of bias direction plots (via the tool described in Appendix \@ref(appendix-robvis)), calculation of bias/indirectness-adjusted estimates for each result and synthesis of these adjusted results in a standard random-effects meta-analysis.

@@

The second key output is an annotated example dataset for the LDL-c/Alzheimer's disease causal question addressed in this chapter. Available via the `triangulate` package, this data will provide developers of new triangulation methods with an example dataset on which to test new tools, something which seriously hindered development of this work.

>>>>

```{block type='savequote', include=knitr::is_latex_output(), quote_author='(ref:tri-quote)', echo = TRUE}
Next on my list of features to be specially considered I would place the consistency of the observed association. Has it been repeatedly observed by different persons, in different places, circumstances and times?
```
(ref:tri-quote) --- Austin Bradford Hill, 1965[@hill1965]

@@

There are two key outputs from this chapter. The first is an R package, `triangulate`, available at [https://github.com/mcguinlu/triangulate](https://github.com/mcguinlu/triangulate). Personal communication with the authors of the bias-/indirectness-adjusted meta-analysis approach resulted in the STATA code to implement the adjusted model.[@turner2009] This has since been generalised as part of the `triangulate` R package to enable other users to apply the approach detailed here. The package allows for preprocessing of domain-based risk-of-bias data to correctly identify the direction of bias relative to the effect estimate, use of domain-specific prior distributions, production of bias direction plots (via the tool described in Appendix \@ref(appendix-robvis)), calculation of bias/indirectness-adjusted estimates for each result and synthesis of these adjusted results in a standard random-effects meta-analysis.

@@

The second key output is an example dataset for the LDL-c/Alzheimer's disease causal question addressed in this chapter. Available via the `triangulate` package, this data will provide developers of new triangulation methods with an example dataset on which to test new tools, something which seriously hindered development of this work.

### 2021-12-17 08-Discussion.Rmd
Submitted version

::: {.laybox data-latex=""}
## Lay summary

@@

The first is a large-scale analysis of 953,635 patients in the CPRD, followed from their first LDL-c measurement.[@iwagami2021] The study found evidence for a slightly increased risk of dementia with higher LDL-c levels at mid-life, driven by the finding for the Alzheimer's disease subgroup. These results are consistent with the findings presented in this thesis, particularly as this study did not have access to _Apo_$\mathcal{E}4$ status and so could not adjust for confounding by this variable. _Apo_$\mathcal{E}4$ is a key risk factor for both Alzheimer's disease and raised LDL-c levels, <!-- TODO CITATION NEEDED --> and so could explain the harmful association observed. This is supported by a simulation analysis reported within this study, which used published information to predict that _Apo_$\mathcal{E}4$ would induce a risk ratio of 1.09 per 1-SD increase in LDL-c in the absence of a true causal effect.

>>>>

::: {.laybox data-latex=""}
## Lay summary {-}

@@

The first is a large-scale analysis of 953,635 patients in the CPRD, followed from their first LDL-c measurement.[@iwagami2021] The study found evidence for a slightly increased risk of dementia with higher LDL-c levels at mid-life, driven by the finding for the Alzheimer's disease subgroup. These results are consistent with the findings presented in this thesis, particularly as this study did not have access to _Apo_$\mathcal{E}4$ status and so could not adjust for confounding by this variable. _Apo_$\mathcal{E}4$ is a key risk factor for both Alzheimer's disease and raised LDL-c levels,[@bennet2007] and so could explain the harmful association observed. This is supported by a simulation analysis reported within this study, which used published information to predict that _Apo_$\mathcal{E}4$ would induce a risk ratio of 1.09 per 1-SD increase in LDL-c in the absence of a true causal effect.

### 2021-12-17 front-and-back-matter/99-appendices.Rmd
Submitted version

In response to the COVID-19 pandemic, I worked with a team of evidence synthesists to develop a literature searching/screening pipeline. This work underpinned a living systematic review on impact of the COVID-19 pandemic on self-harm and suicidal behaviour. Subsequently, the pipeline enabled rapid reviews of related topics, such as an analysis of the available evidence by population.

@@

A copy of the risk of bias tool used to assess Mendelian randomisation studies is presented in Table \@ref(tab:mrTool-table), below.

@@

Enforcing a minimum period of follow-up is normally used to reduce the risk of reverse causation in an analysis by excluding events occurring within a certain time-frame post-exposure. With respect to dementia outcomes, which have a long prodromal period, enforcing a minimum period of follow-up can add biological validity to an analysis. Incident cases of dementia that occur within a few years of starting a drug may already be beyond effective preventive treatment at the time the drug was started, and so restricting the analysis to those with longer exposure to treatment may provide a more reliable estimate of the effect.

@@

However, applying a similar approach to the analysis performed in Chapter \@ref(cprd-analysis-heading), illustrated in Panel B of Figure \@ref(fig:minFU), is challenging. The use of a time updating treatment indicator results in two potential starting points for the 10-year lag: entry into the cohort or initiation of statin therapy. As the purpose of this analysis is to investigate the impact of statin treatment, starting the minimum period of follow-up from cohort entry is not helpful as it does not ensure that participants have been exposed to the drug for 10 years. Alternatively, defining the minimum period of follow-up from statin initiation cause issues around the treatment of time-at-risk before statin initiation and around the start point for the follow-up period in those never prescribed a statin. In summary, either approach is likely to be substantially biased.

@@

(ref:fpLdlAdComparison-cap) __Comparison of bias-/indirectness-adjusted meta-analyses of LDL-c/Alzheimer's disease under two different scenarios of additive bias.__ Scenario 1: position value between the additive distributions assigned to "High" and "Moderate" bias levels is the same as between the "Moderate" and "Low" levels. Scenario 2: position value between the additive distributions assigned to "High" and "Moderate" bias levels is twice that between "Moderate" and "Low" levels. See Section \@ref(tri-bias-assign) for more details.

@@

(ref:fpTgVaDComparison-cap) __Comparison of bias-/indirectness-adjusted meta-analyses of triglycerides/vascular dementia under two different scenarios of prior distributions of bias.__ Scenario 1: position value between the additive distributions assigned to "High" and "Moderate" bias levels is the same as between the "Moderate" and "Low" levels. Scenario 2: position value between the additive distributions assigned to "High" and "Moderate" bias levels is twice that between "Moderate" and "Low" levels. See Section \@ref(tri-bias-assign) for more details.

@@

Reproducible and science has been a key theme running through this thesis, as reflected by the development of an open-source tools to search preprinted literature, create visualisations of risk-of-bias data and to preform quantitative triangulation. In line with this, an open-source copy of the code used to produce this thesis is available on GitHub, as is the code used to perform the analysis contained within it.

@@

As part of a reflective process, during the course of my PhD I kept a record of things (analytical mishaps, failed experiments, and unsuccessful grant applications) I attempted but which did not work. These are summarised briefly here:

@@

<!-- TODO Need to complete this sections -->
The `robvis` package provides functions to visualise risk-of-bias assessments, with the resulting plots formatted based on the specific risk-of-bias assessment tool used. The tool is documented extensively online at [https://mcguinlu.github.io/robvis/](https://mcguinlu.github.io/robvis/). As such, it is not described in detail here, other than to illustrate the data structures required for the figures presented in this thesis and to describe its reception in the evidence synthesis community.

@@

For the paired forest plots presented in Chapter \@ref(sys-rev-results-heading), the tool expects a "Study" column, followed by one column per domain of bias and one column for the overall risk of bias judgement. To illustrate this, the `data_rob2` dataset which is available via the `robvis` package and contains example risk-of-bias assessments performed using the RoB2.0 tool for randomized controlled trials, is shown in Table \@ref(tab:biasData-table).

>>>>

In response to the COVID-19 pandemic, I worked with a team of evidence synthesists to develop a literature searching/screening pipeline. This work underpinned a living systematic review on impact of the COVID-19 pandemic on self-harm and suicidal behaviour. Subsequently, the pipeline enabled rapid reviews of related questions, such as the impact on specific study populations.

@@

A copy of the risk-of-bias tool used to assess Mendelian randomisation studies is presented in Table \@ref(tab:mrTool-table), below.

@@

Enforcing a minimum period of follow-up is normally used to reduce the risk of reverse causation in an analysis by excluding events occurring within a certain time-frame post exposure. With respect to dementia outcomes, which have a long prodromal period, enforcing a minimum period of follow-up can add biological validity to an analysis. Incident cases of dementia that occur within a few years of starting a drug may already be beyond effective preventive treatment at the time the drug was started, and so restricting the analysis to those with longer exposure to treatment may provide a more reliable estimate of the effect.

@@

However, applying a similar approach to the analysis performed in Chapter \@ref(cprd-analysis-heading), illustrated in Panel B of Figure \@ref(fig:minFU), is challenging. The use of a time updating treatment indicator results in two potential starting points for the 10-year lag: entry into the cohort or initiation of statin therapy. As the purpose of this analysis is to investigate the impact of statin treatment, starting the minimum period of follow-up from cohort entry is not helpful as it does not ensure that participants have been exposed to the drug for 10 years. Alternatively, defining the minimum period of follow-up from statin initiation cause issues around the treatment of time-at-risk before statin initiation and around the start point for the follow-up period in those never prescribed a statin. In summary, either approach is likely to be substantially biased, and so an analysis using a minimum follow-up period was not performed.

@@

(ref:fpLdlAdComparison-cap) __Comparison of bias-/indirectness-adjusted meta-analyses of LDL-c/Alzheimer's disease under two different scenarios of additive bias__ - Scenario 1: position value between the additive distributions assigned to "High" and "Moderate" bias levels is the same as between the "Moderate" and "Low" levels. Scenario 2: position value between the additive distributions assigned to "High" and "Moderate" bias levels is twice that between "Moderate" and "Low" levels. See Section \@ref(tri-bias-assign) for more details.

@@

(ref:fpTgVaDComparison-cap) __Comparison of bias-/indirectness-adjusted meta-analyses of triglycerides/vascular dementia under two different scenarios of prior distributions of bias__ - Scenario 1: position value between the additive distributions assigned to "High" and "Moderate" bias levels is the same as between the "Moderate" and "Low" levels. Scenario 2: position value between the additive distributions assigned to "High" and "Moderate" bias levels is twice that between "Moderate" and "Low" levels. See Section \@ref(tri-bias-assign) for more details.

@@

Reproducible research has been a key theme running through this thesis, as reflected by the development of an open-source tools to search preprinted literature, create visualisations of risk-of-bias data and to preform quantitative triangulation. In line with this, an open-source copy of the code used to produce this thesis is available on GitHub, as is the code used to perform the analysis contained within it.

@@

As part of a reflective process, I kept a record of things (analytical mishaps, failed experiments, and unsuccessful grant applications) I attempted during the course of my PhD but which did not work. These are summarised briefly here:

@@

The `robvis` package provides functions to visualise risk-of-bias assessments, with the resulting plots formatted based on the specific risk-of-bias assessment tool used. The tool is documented extensively online at [https://mcguinlu.github.io/robvis/](https://mcguinlu.github.io/robvis/). As such, it is not described in detail here, other than to illustrate the data structures required for the figures presented in this thesis and to describe its reception in the evidence synthesis community.

@@

For the paired forest plots presented in Chapter \@ref(sys-rev-results-heading), the tool expects a "Study" column, followed by one column per domain of bias and one column for the overall risk of bias judgement. To illustrate this, the `data_rob2` dataset which is available via the `robvis` package and contains example risk-of-bias assessments performed using the RoB2.0 tool for randomised controlled trials, is shown in Table \@ref(tab:biasData-table).

### 2021-12-17 front-and-back-matter/_abstract.Rmd
Submitted version

__Results:__ The systematic review did not identify a consistent effect of blood lipids on any dementia outcome across study designs considered, though there was some suggestion of a protective effect of LDL-c lowering on all-cause dementia and Alzheimer's disease in observational studies of statin use. Analysis of the CPRD data provided weak evidence for a protective effect of statins on all-cause dementia and Alzheimer's disease but suggested a harmful association with vascular dementia,  likely due to confounding by indication. The IPD meta-analysis provided some evidence for the association of triglycerides and vascular dementia only. Finally, the quantitative triangulation analysis did not provide strong evidence for the causal effect of blood lipids on dementia outcomes.

>>>>

__Results:__ The systematic review did not identify a consistent effect of blood lipids on any dementia outcome across the study designs considered, though there was some suggestion of a protective effect of LDL-c lowering on all-cause dementia and Alzheimer's disease in observational studies of statin use. Analysis of the CPRD data provided weak evidence for a protective effect of statins on all-cause dementia and Alzheimer's disease but suggested a harmful association with vascular dementia,  likely due to confounding by indication. The IPD meta-analysis provided some evidence for the association of triglycerides and vascular dementia only. Finally, the quantitative triangulation analysis did not provide strong evidence for the causal effect of blood lipids on dementia outcomes.

### 2021-12-17 front-and-back-matter/_acknowledgements.Rmd
Submitted version

First and foremost, I must thank my primary supervisor, Julian Higgins, for his tireless support and encouragement. I'd also like to thank my other supervisors - Yoav Ben-Shlomo, Liz Coulthard and George Davey-Smith - for all their input over the past three years. Thanks also to Rupert Payne and Emma Anderson for their engaging discussion and useful feedback during my annual progress reviews.

@@

Finally, a big thanks to Vivienne Gardiner, who sacrificed a substantial amount of time to proof-reading this thesis.

@@

\begin{flushright}
Canynge Hall, Bristol \\\
1 December 2021
\end{flushright}

>>>>

First and foremost, I must thank my primary supervisor, Julian Higgins, for his tireless support and encouragement. I'd also like to thank my other supervisors - Yoav Ben-Shlomo, Liz Coulthard and George Davey Smith - for all their input over the past three years. Thanks also to Rupert Payne and Emma Anderson for their engaging discussion and useful feedback during my annual progress reviews.

@@

Finally, a big thanks to Vivienne Gardiner, who sacrificed a substantial amount of time to proof-read this thesis.

@@

\begin{flushright}
Canynge Hall, Bristol \\\
17 December 2021
\end{flushright}
